Carbonate chemistry at Ny-Ålesund

### 2021-12-09 awipev-CO2.Rmd
Work on tratification

d_all <- d_hour %>%
  dplyr::mutate_all(function(x) ifelse(is.nan(x), NA, x)) %>% # replace all NaN by NA
  dplyr::mutate(depth_group = dplyr::case_when(depth > 0 & depth <= 2 ~ '0-2',
                depth > 2 & depth <= 4 ~ '2-4',
                depth > 4 & depth <= 6 ~ '4-6',
                depth > 6 & depth <= 8 ~ '6-8',
                depth > 8 & depth <= 12 ~ '>8', #8 to 12.4
                TRUE ~ 'NA'
                ),
                sal_insitu_ctd_1m = ifelse(depth_group == '0-2', sal_insitu_ctd, NA),
                sal_insitu_ctd_3m = ifelse(depth_group == '2-4', sal_insitu_ctd, NA),
                sal_insitu_ctd_5m = ifelse(depth_group == '4-6', sal_insitu_ctd, NA),
                sal_insitu_ctd_7m = ifelse(depth_group == '6-8', sal_insitu_ctd, NA),
                sal_insitu_ctd_10m = ifelse(depth_group == '>8', sal_insitu_ctd, NA),
                temp_insitu_ctd_1m = ifelse(depth_group == '0-2', temp_insitu_ctd, NA),
                temp_insitu_ctd_3m = ifelse(depth_group == '2-4', temp_insitu_ctd, NA),
                temp_insitu_ctd_5m = ifelse(depth_group == '4-6', temp_insitu_ctd, NA),
                temp_insitu_ctd_7m = ifelse(depth_group == '6-8', temp_insitu_ctd, NA),
                temp_insitu_ctd_10m = ifelse(depth_group == '>8', temp_insitu_ctd, NA),
  ) %>% 
  dplyr::rowwise() %>% 
  dplyr::mutate(n_depth_ctd = sum(is.na( # create new variable to identify the hours for which there are CTD data at more than 1 depth 
  c(
    temp_insitu_ctd_1m,
    temp_insitu_ctd_3m,
    temp_insitu_ctd_5m,
    temp_insitu_ctd_7m,
    temp_insitu_ctd_10m
  )
), na.rm = TRUE)) %>% 
  ungroup()

@@

# we look at 0 to 4 m vs more than 8 m
tmp <- d_all %>% 
  dplyr::select(montht, pHint_tot_sf, depth) %>%
  dplyr::mutate(depth = as.factor(ifelse(depth == 1 | depth == 3, 2, depth))) %>% 
  dplyr::filter(depth == 2 | depth == 10) %>% 
  dplyr::group_by(depth, montht)
summary(aov(tmp$pHint_tot_sf ~ tmp$depth * tmp$montht))

@@

## Stratification
```{r stratification, echo=FALSE, message= FALSE, warning=FALSE}
# # we look at the distribution of data
# tmp <- d_all %>% 
#   dplyr::select(date, depth, pressure_insitu_ctd) %>% 
#   dplyr::mutate(month=month(d_all$date, label = TRUE)) %>% 
#   group_by(month)
depth_labels <- c("0-2", "2-4", "4-6", "6-8", ">8")
# we look at salinity all depths
tmp <- d_all %>% 
  #dplyr::mutate(depth_group = depth_group_pressure) %>% 
  dplyr::select(montht, sal_insitu_ctd, depth, depth_group) %>%
  tidyr::drop_na() %>% 
  #dplyr::filter(!is.na(sal_insitu_ctd), !is.na(depth_group)) %>% 
  group_by(depth_group, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = sal_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Salinity vs depth", x = "Depth (m)", y= "Salinity") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)

summary(aov(tmp$sal_insitu_ctd ~ tmp$depth_group * tmp$montht))

# we look at salinity 0 to 4 m vs more than 8 m
tmp <- d_all %>% 
  dplyr::mutate(depth_group = depth_group_pressure) %>% 
  dplyr::select(montht, sal_insitu_ctd, depth_group) %>%
  dplyr::mutate(depth = as.factor(ifelse(depth_group == 1 | depth_group == 3, 2, depth_group))) %>% 
  dplyr::filter(depth == 2 | depth == 10)
tmp_delta <- tmp %>% 
  group_by(depth, montht) %>% 
  dplyr::summarise(sal_med = median(sal_insitu_ctd, na.rm = TRUE)) %>% 
  tidyr::pivot_wider(names_from = depth, values_from = sal_med) %>% 
  dplyr::rename(median0_4m = "2",
                median8_12m = "10") %>% 
  dplyr::mutate(sal_delta = median0_4m - median8_12m)
kable(tmp_delta, digits = 2, caption = "Median salinity with (median0_4m - median8_12m)")

>>>>

d_all <- d_hour %>%
  dplyr::mutate_all(function(x) ifelse(is.nan(x), NA, x)) %>% # replace all NaN by NA
  dplyr::mutate(depth_group = as.factor(dplyr::case_when(depth > 0 & depth <= 2 ~ '0-2',
                depth > 2 & depth <= 4 ~ '2-4',
                depth > 4 & depth <= 6 ~ '4-6',
                depth > 6 & depth <= 8 ~ '6-8',
                depth > 8 & depth <= 12 ~ '>8', #8 to 12.4
                TRUE ~ 'NA'
                )),
                sal_insitu_ctd_1m = ifelse(depth_group == '0-2', sal_insitu_ctd, NA),
                sal_insitu_ctd_3m = ifelse(depth_group == '2-4', sal_insitu_ctd, NA),
                sal_insitu_ctd_5m = ifelse(depth_group == '4-6', sal_insitu_ctd, NA),
                sal_insitu_ctd_7m = ifelse(depth_group == '6-8', sal_insitu_ctd, NA),
                sal_insitu_ctd_10m = ifelse(depth_group == '>8', sal_insitu_ctd, NA),
                temp_insitu_ctd_1m = ifelse(depth_group == '0-2', temp_insitu_ctd, NA),
                temp_insitu_ctd_3m = ifelse(depth_group == '2-4', temp_insitu_ctd, NA),
                temp_insitu_ctd_5m = ifelse(depth_group == '4-6', temp_insitu_ctd, NA),
                temp_insitu_ctd_7m = ifelse(depth_group == '6-8', temp_insitu_ctd, NA),
                temp_insitu_ctd_10m = ifelse(depth_group == '>8', temp_insitu_ctd, NA),
  ) %>% 
  dplyr::rowwise() %>% 
  dplyr::mutate(n_depth_ctd = sum(is.na( # create new variable to identify the hours for which there are CTD data at more than 1 depth 
  c(
    temp_insitu_ctd_1m,
    temp_insitu_ctd_3m,
    temp_insitu_ctd_5m,
    temp_insitu_ctd_7m,
    temp_insitu_ctd_10m
  )
), na.rm = TRUE)) %>% 
  ungroup()

@@

# we look at 0 to 4 m vs more than 8 m
tmp <- d_all %>% 
  dplyr::select(montht, pHint_tot_sf, depth) %>%
  dplyr::mutate(depth2 = as.factor(case_when(depth < 4 ~ '0-4', 
                                             depth > 8~ '8-12',
                                             TRUE ~ NA_real_))) %>% 
  dplyr::filter(depth2 == '0-2' | depth2 == '8-12') %>% 
  dplyr::group_by(depth2, montht)
summary(aov(tmp$pHint_tot_sf ~ tmp$depth2 * tmp$montht))

@@

## Stratification
```{r stratification, echo=FALSE, message= FALSE, warning=FALSE}
# # we look at the distribution of data
depth_labels <- c("0-2", "2-4", "4-6", "6-8", ">8")
# we look at salinity all depths
tmp <- d_all %>% 
  dplyr::select(montht, sal_insitu_ctd, depth, depth_group) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  group_by(depth_group, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = sal_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Salinity vs depth", x = "Depth (m)", y= "Salinity") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$sal_insitu_ctd ~ tmp$depth_group * tmp$montht))

# we look at salinity 0 to 4 m vs more than 8 m
tmp <- d_all %>% 
  dplyr::mutate(depth_group = depth_group_pressure) %>% 
  dplyr::select(montht, sal_insitu_ctd, depth_group) %>%
  dplyr::mutate(depth2 = as.factor(case_when(depth < 4 ~ '0-4', 
                                             depth > 8~ '8-12',
                                             TRUE ~ NA_real_))) %>% 
  dplyr::filter(depth2 == '0-4' | depth2 == '8-12') %>% 
  dplyr::group_by(depth2, montht)
tmp_delta <- tmp %>% 
  group_by(depth, montht) %>% 
  dplyr::summarise(sal_med = median(sal_insitu_ctd, na.rm = TRUE)) %>% 
  tidyr::pivot_wider(names_from = depth, values_from = sal_med) %>% 
  dplyr::rename(median0_4m = "2",
                median8_12m = "10") %>% 
  dplyr::mutate(sal_delta = median0_4m - median8_12m)
kable(tmp_delta, digits = 2, caption = "Median salinity with (median0_4m - median8_12m)")
